### 2015-12-10 paper/2-background.tex

Deadlock can occur when a dependency cycle occurs and the amount of exchanged messages is low.
Again, multiple algorithms are defined to handle this situation.
Algorithms range from deadlock avoidance to deadlock recovery.

The main advantage of this approach is that it has a very low overhead in cases of high parallelism between different nodes.
Each node can simulate in parallel, while sporadically notifying other nodes that they can progress even further.
The disadvantage, however, is that the amount of parallelism is explicitly limited by the size of the lookahead.
If a node can influence another (almost) instantaneously, no matter how rarely it occurs, the amount of parallelism is severely reduced.
It is also up to the user to define the size of the lookahead, depending on how the model is known to behave.
Slight changes in model behaviour can cause significant changes to the lookahead, and can therefore also have a significant influence on simulation performance.

@@

Further changes are required when the model performs unrecoverable transactions, such as I/O (\textit{e.g.}, tracing, writing to file, printing output), allocating and deallocating memory, and so on.
For this, lightweight synchronization algorithms are still required to determine the lower bound of all simulation times, through the computation of a \textit{Global Virtual Time} (GVT).

Optimistic synchronization never blocks: a node always progresses as fast as possible, at the cost of potentially discarding simulation results.
More importantly, performance is not limited by a small lookahead, caused by a very infrequent event.
If an (almost) instantaneous event occurs rarely, performance will only be impacted if it occurs, and not at every simulation step.
The main disadvantage is unpredictable performance and arbitrary cost of rollbacks due to the propagation of causality errors.
If rollbacks occur frequently, simulation quickly becomes slow, as the overhead of the recovery mechanisms becomes significant.
Apart from overhead in CPU time, a significant memory overhead is present: all intermediate states qre stored up to a point where it can be considered \textit{irreversible}.

>>>>

Deadlock can occur when a dependency cycle occurs and the amount of exchanged messages is low.
Multiple algorithms are defined to handle this situation, such as deadlock avoidance and deadlock recovery.

The main advantage of conservative synchronization is that it has a low overhead if high parallelism exists between nodes.
Each node can simulate in parallel, while sporadically notifying other nodes that they can progress even further.
The disadvantage, however, is that the amount of parallelism is explicitly limited by the size of the lookahead.
If a node can influence another (almost) instantaneously, no matter how rarely it occurs, the amount of parallelism is severely reduced.
It is also up to the user to define the size of the lookahead, depending on how the model is known to behave.
Slight changes in model behaviour can cause significant changes to the lookahead, and can therefore also have a significant influence on simulation performance.

@@

Further changes are required for unrecoverable operations, such as I/O (\textit{e.g.}, tracing, writing to file, printing output) and memory management.
Lightweight algorithms are still required to determine the lower bound of all simulation times, through the computation of a \textit{Global Virtual Time} (GVT).

The main advantage is that performance is not limited by a small lookahead, caused by a very infrequent event.
If an (almost) instantaneous event occurs rarely, performance will only be impacted if it occurs, and not at every simulation step.
The main disadvantage is unpredictable performance and arbitrary cost of rollbacks due to the propagation of causality errors.
If rollbacks occur frequently, simulation quickly becomes slow, as the overhead of the recovery mechanisms becomes significant.
Apart from overhead in CPU time, a significant memory overhead is present: all intermediate states qre stored up to a point where it can be considered \textit{irreversible}.

### 2015-12-10 paper/4-performance.tex

Figure~\ref{fig:phold_priority} shows how simulation performance is influenced by the fraction of high-priority events.
If barely any high-priority events occur, conservative synchronization is penalized due to its excessive blocking, which often turned out to be unnecessary.
Should many high-priority events occur, however, optimistic synchronization is penalized due to its mindless progression of simulation, which frequently needed to be reverted.

>>>>

Figure~\ref{fig:phold_priority} shows how simulation performance is influenced by the fraction of high-priority events.
If barely any high-priority events occur, conservative synchronization is penalized due to its excessive blocking, which often turned out to be unnecessary.
Should many high-priority events occur, however, optimistic synchronization is penalized due to its mindless progression of simulation, which frequently needed to be reverted.
These results show that there is no single perfect synchronization algorithm for this model.
Depending on model configuration, either synchronization protocol might be better.
We therefore claim that our contribution is invaluable for high performance simulation: depending on the observed communication behaviour, modellers can use the most appropriate synchronization protocol.
